Innovations Report: Procedural Fur System

### 2018-03-19 docs/paper.tex
docs tweaks

\subsection{Related work} \label{sec:related}
Whilst many larger companies opt to develop their own software for handling fur, such as MPC's Furtility~\cite{furtility}, there are a number of commercially available solutions that are also widely used in production.

The most popular of these is XGen~\cite{xgen}, originally developed by Walt Disney Animation Studios, which has been included in Autodesk's Maya since 2014. XGen uses a 'Collection' and 'Description' paradigm; Descriptions store user controlled parameters which affect fur primitives (curves for example) on a select area of a mesh, and Collections act as a container for these Descriptions. There are two primary methods offered for distributing primitives on a mesh, Guide Curves can be used to manually control the placement and shape of primitives by placing 'Guides' on the mesh surface, whereas 'Groomable Splines' randomly distributes splines over the mesh surface which can be manipulated in real time using viewport manipulators that mimic the brushing and styling of fur. In both modes, 'modifiers' can be added to the modifier stack to further adjust the visual outcome, adding features such as clumping, noise, and coiling. Attributes within these modifiers can be further controlled using expressions or texture maps.

Another commercially available solution is Yeti~\cite{yeti}, which is also offered as a plugin within Autodesk's Maya. Yeti offers a more procedural approach, giving the user a node graph interface in which to construct the desired look from individual components. For example, the most simple usage requires the user to first create a 'scatter' node to distribute points on the mesh surface, and then combine it with a 'grow' node to extrude curves from these points. Whilst at first this may seem unintuitive when compared to XGen, it offers greater flexibility and avoids the pitfalls associated with a destructive workflow.

@@

\subsection{GPU acceleration} \label{sec:existingGPU}
Whilst it is not clear to what extent each of these solutions make use of GPU acceleration, they all perform fast enough to allow the user to manipulate them interactively. However, software such as the HairWorks \cite{hairworks} API and TressFX \cite{tressfx} library are developed primarily for use in games, so they are designed to utilise the GPU as much as possible. These solutions are able to perform both simulation and rendering of hair in real-time, providing interactivity that exceeds that of comparable software designed for VFX usage. It is worth noting that these tools do not provide a method for authoring hair and fur content, but instead designed to use load content generated by 3rd party authoring tools.

@@

I wanted to include a node-graph style interface within my application, as these are commonly used within existing artist tools (Autodesk Maya, Unreal Engine), and would encourage modularity within my API. Qt does not natively provide this kind of interface, so I made use of NodeEditor~\cite{Pinaev2017}, an existing Qt-based library that provides this functionality.

@@

Another desirable addition to my system would be viewport manipulators, similar to the 'Groomable Splines' feature in XGen. This could allow me to implement techniques such as those described in Skippy~\cite{skippy}, which would provide my system with a unique user experience compared to commercially available systems.

>>>>

\subsection{Related work} \label{sec:related}
Whilst many larger companies opt to develop their own software for handling fur, such as MPC's \textit{Furtility}~\cite{furtility}, there are a number of commercially available solutions that are also widely used in production.

The most popular of these is \textit{XGen}~\cite{xgen}, originally developed by Walt Disney Animation Studios, which has been included in Autodesk's Maya since 2014. XGen uses a 'Collection' and 'Description' paradigm; Descriptions store user controlled parameters which affect fur primitives (curves for example) on a select area of a mesh, and Collections act as a container for these Descriptions. There are two primary methods offered for distributing primitives on a mesh, Guide Curves can be used to manually control the placement and shape of primitives by placing 'Guides' on the mesh surface, whereas 'Groomable Splines' randomly distributes splines over the mesh surface which can be manipulated in real time using viewport manipulators that mimic the brushing and styling of fur. In both modes, 'modifiers' can be added to the modifier stack to further adjust the visual outcome, adding features such as clumping, noise, and coiling. Attributes within these modifiers can be further controlled using expressions or texture maps.

Another commercially available solution is \textit{Yeti}~\cite{yeti}, which is also offered as a plugin within Autodesk's Maya. Yeti offers a more procedural approach, giving the user a node graph interface in which to construct the desired look from individual components. For example, the most simple usage requires the user to first create a 'scatter' node to distribute points on the mesh surface, and then combine it with a 'grow' node to extrude curves from these points. Whilst at first this may seem unintuitive when compared to XGen, it offers greater flexibility and avoids the pitfalls associated with a destructive workflow.

@@

\subsection{GPU acceleration} \label{sec:existingGPU}
Whilst it is not clear to what extent each of these solutions make use of GPU acceleration, they all perform fast enough to allow the user to manipulate them interactively. However, software such as the \textit{HairWorks}~\cite{hairworks} API and \textit{TressFX}~\cite{tressfx} library are developed primarily for use in games, so they are designed to utilise the GPU as much as possible. These solutions are able to perform both simulation and rendering of hair in real-time, providing interactivity that exceeds that of comparable software designed for VFX usage. It is worth noting that these tools do not provide a method for authoring hair and fur content, but instead designed to use load content generated by 3rd party authoring tools.

@@

I wanted to include a node-graph style interface within my application, as these are commonly used within existing artist tools (Autodesk Maya, Unreal Engine), and would encourage modularity within my API. Qt does not natively provide this kind of interface, so I made use of \textit{NodeEditor}~\cite{Pinaev2017}, an existing Qt-based library that provides this functionality.

@@

Another desirable addition to my system would be viewport manipulators, similar to the 'Groomable Splines' feature in XGen. This could allow me to implement techniques such as those described in \textit{Skippy}~\cite{skippy}, which would provide my system with a unique user experience compared to commercially available systems.
